An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-04-14 EconGrowth-R-Functions.Rnw
Added the option to include CES models in partialResidualPlots.

#############################
# This function creates a graph for Cobb-Douglas with energy functions. The graph has 
# a double Y axis. The first Y axis has alpha, beta, and gamma vs. gamma. The second Y axis
# has SSE vs. gamma.
# countryAbbrev tells which country you want to view
# energyType tells which type of energy you want
# textScaling tells how you want to scale the text on the axes and other places.
# This 
##
cdeGridGraph <- function(countryAbbrev, energyType, keyXLoc = defaultKeyXLoc, keyYLoc = defaultKeyYLoc, textScaling = 1.0){
  gammaGrid <- seq(0.0, 0.95, by=0.05)
  data <- cdeGridData(countryAbbrev, energyType, gammaGrid)
  # Now melt the data so that it is in the correct form.
  data_molten <- melt(data, id="gamma")
  alphaBetaGammaRange <- c(0.0, 1.0)
  alphaBetaGammaTics <- c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)
  lambdaRange <- c(0.0, 0.004)
  lambdaTics <- c(0.0, 0.001, 0.002, 0.003, 0.004)
  sseRange <- c(0.0, 0.08)
  sseTics <- c(0.0, 0.02, 0.04, 0.06, 0.08)
  # Create a lattice graph that shows the four variables (lambda, alpha, beta, and SSE) vs. gamma.
  graphType <- "b" #b is for both symbls and lines. Necessary so that the key looks right.
  lineTypes <- c(2, 1, 3, 0) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(1, 3, 1, 0) #line widths. 0 means no line.
  symbols <- c(NA, NA, NA, 1)
  colors <- c("black", "black", "black", "black") #line and symbol colors
  abgGraph <- xyplot(alpha+beta+gamma ~ gamma, data=data,
                     type=graphType,
                     lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                     pch = symbols, col.symbol = colors, #Controls symbol parameters
                     key=list(text=list(c("$\\alpha$", "$\\beta$", "$\\gamma$", "$SSE_y$")),
                              type=graphType,
                              cex=keyTextSize * textScaling, #controls size of text in the key
                              lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                              pch=symbols, col=colors, #controls symbol (plot characters) types
                              columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                     scales=list(cex=scaleTextSize * textScaling, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 x=list(at=c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)), #location for x tics
                                 y=list(at=c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0))  #location for y tics
                     ), 
                     ylim=c(0.0, 1.5), #y axis limits
                     #axis labels and scaling
                     xlab=list(label="$\\gamma$", cex=textScaling), 
                     ylab=list(label="$\\alpha$, $\\beta$, $\\gamma$", cex=textScaling)
  )
  sseGraph <- xyplot(SSE ~ gamma, data=data,
                     pch = c(1), col.symbol = c("black"), #Controls symbol parameters
                     scales=list(cex=scaleTextSize * textScaling, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 y=list(at=c(0.0, 0.02, 0.04, 0.06, 0.08)) #location for y tics                                          
                                 ), 
                     ylim=c(0.0, 0.08),
                     xlab=list(label="$\\gamma$", cex=textScaling), #Not used by the double-Y graph.
                     ylab=list(label="$SSE$", cex=textScaling)
                     )
  # Combine the two graphs into a double-Y graph.
  doubleYGraph <- doubleYScale(abgGraph, sseGraph, add.ylab2=TRUE, use.style=FALSE)
  return(doubleYGraph)
}

@@

#################
# Creates a plot with y residuals vs. energy (e). The y residuals are from a Cobb-Douglas model without energy.
# The energy term on the x axis is whatever type is specified in the calling funciton
partialResidualPlot <- function(countryAbbrev, energyType, textScaling=1.0){
  # Get the model
  cdModel <- cdModel(countryAbbrev)
  # Get the residuals
  cdResid <- resid(cdModel)
  cdResid <- data.frame(cdResid)
  cdResid <- padRows(countryAbbrev=countryAbbrev, df=cdResid)
  # Load data
  data <- loadData(countryAbbrev)
  data <- cbind(data, cdResid)
  data <- replaceColName(dataTable=data, factor=energyType, newName="iEToFit")
  plot <- xyplot(cdResid ~ iEToFit, data=data,
#                 type=c("p",:smooth"),
                 type=c("p"),
                 col="black",
                 scales=list(cex=scaleTextSize*textScaling, #controls text size on scales.
                             tck=scaleTickSize), #controls tick mark length. < 0 for inside the graph.
                 xlab=list(label=paste("$", tolower(energyType), "$", sep=""), cex=textScaling),
                 ylab=list(label="$y$ residuals (CD without energy)", cex=textScaling)
                 )
  return(plot)  
}
@

@@

>>>>

#############################
# This function creates a graph for Cobb-Douglas with energy functions. The graph has 
# a double Y axis. The first Y axis has alpha, beta, and gamma vs. gamma. The second Y axis
# has SSE_y vs. gamma.
# countryAbbrev tells which country you want to view
# energyType tells which type of energy you want
# textScaling tells how you want to scale the text on the axes and other places.
##
cdeGridGraph <- function(countryAbbrev, energyType, keyXLoc = defaultKeyXLoc, keyYLoc = defaultKeyYLoc, textScaling = 1.0){
  gammaGrid <- seq(0.0, 0.95, by=0.05)
  data <- cdeGridData(countryAbbrev, energyType, gammaGrid)
  # Now melt the data so that it is in the correct form.
  data_molten <- melt(data, id="gamma")
  alphaBetaGammaRange <- c(0.0, 1.0)
  alphaBetaGammaTics <- c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)
  lambdaRange <- c(0.0, 0.004)
  lambdaTics <- c(0.0, 0.001, 0.002, 0.003, 0.004)
  sseRange <- c(0.0, 0.08)
  sseTics <- c(0.0, 0.02, 0.04, 0.06, 0.08)
  # Create a lattice graph that shows the four variables (lambda, alpha, beta, and SSE) vs. gamma.
  graphType <- "b" #b is for both symbls and lines. Necessary so that the key looks right.
  lineTypes <- c(2, 1, 3, 0) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(1, 3, 1, 0) #line widths. 0 means no line.
  symbols <- c(NA, NA, NA, 1)
  colors <- c("black", "black", "black", "black") #line and symbol colors
  abgGraph <- xyplot(alpha+beta+gamma ~ gamma, data=data,
                     type=graphType,
                     lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                     pch = symbols, col.symbol = colors, #Controls symbol parameters
                     key=list(text=list(c("$\\alpha$", "$\\beta$", "$\\gamma$", "$SSE_y$")),
                              type=graphType,
                              cex=keyTextSize * textScaling, #controls size of text in the key
                              lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                              pch=symbols, col=colors, #controls symbol (plot characters) types
                              columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                     scales=list(cex=scaleTextSize * textScaling, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 x=list(at=c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)), #location for x tics
                                 y=list(at=c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0))  #location for y tics
                     ), 
                     ylim=c(0.0, 1.5), #y axis limits
                     #axis labels and scaling
                     xlab=list(label="$\\gamma$", cex=textScaling), 
                     ylab=list(label="$\\alpha$, $\\beta$, $\\gamma$", cex=textScaling)
  )
  sseGraph <- xyplot(SSE ~ gamma, data=data,
                     pch = c(1), col.symbol = c("black"), #Controls symbol parameters
                     scales=list(cex=scaleTextSize * textScaling, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 y=list(at=c(0.0, 0.02, 0.04, 0.06, 0.08)) #location for y tics                                          
                                 ), 
                     ylim=c(0.0, 0.08),
                     xlab=list(label="$\\gamma$", cex=textScaling), #Not used by the double-Y graph.
                     ylab=list(label="$SSE$", cex=textScaling)
                     )
  # Combine the two graphs into a double-Y graph.
  doubleYGraph <- doubleYScale(abgGraph, sseGraph, add.ylab2=TRUE, use.style=FALSE)
  return(doubleYGraph)
}

@@

@@

<<PartialResidualPlots, eval=TRUE>>=
#################
# Creates a plot with y residuals vs. energy (e). The y residuals are from the specified modelType without energy.
# modelType: one of "CD" or "CES"
# energyType: one of "Q", "X", or "U"
# The energy term on the x axis is whatever type is specified in the calling funciton
# 
partialResidualPlot <- function(modelType, countryAbbrev, energyType, textScaling=1.0){
  if (modelType == "CD"){
    model <- cdModel(countryAbbrev)  
  } else if (modelType == "CES"){
    model <- cesModelNoEnergy(countryAbbrev)
  } else {
    print(paste("Unknown modelType:", modelType, "in partialResidualPlot."))
  }
  # Get the model
  # Get the residuals
  resid <- resid(model)
  resid <- data.frame(resid)
  resid <- padRows(countryAbbrev=countryAbbrev, df=resid)
  # Load data
  data <- loadData(countryAbbrev)
  data <- cbind(data, resid)
  data <- replaceColName(dataTable=data, factor=energyType, newName="iEToFit")
  plot <- xyplot(resid ~ iEToFit, data=data,
                 type=c("p"),
                 col="black",
                 scales=list(cex=scaleTextSize*textScaling, #controls text size on scales.
                             tck=scaleTickSize), #controls tick mark length. < 0 for inside the graph.
                 xlab=list(label=paste("$", tolower(energyType), "$", sep=""), cex=textScaling),
                 ylab=list(label="$y$ residuals (CD without energy)", cex=textScaling)
                 )
  return(plot)  
}

partialResidualLattice <- function(modelType, energyType){

}
@

### 2013-04-14 EconGrowth.Rnw
Added the option to include CES models in partialResidualPlots.

\section{Cobb-Douglas Models With Energy (CDe)}

@@

@@

@@

\subsection{CES without energy}

@@

\subsection{CES with $Q$}

@@

\subsection{CES with $X$}

@@

\subsection{CES with $U$}

>>>>

\section{Cobb-Douglas Models With Energy (CD$e$)}

@@

<<CDPartialResidualsQ, fig.width=5, fig.height=3, fig.align="center", fig.cap="Cobb-Douglas partial residuals plot for thermal energy ($q$).", fig.pos="H", eval=TRUE>>=
partialResidualPlot(modelType="CES", countryAbbrev="US", energyType="Q")
@

@@

<<CDGridSearchGraphs, fig.width=5, fig.height=3, fig.align='center', fig.cap="Cobb-Douglas with thermal energy ($q$) grid search for minimum $SSEy$.", fig.pos="H">>=
cdeGridGraph(countryAbbrev=NA, energyType="Q")
@

@@

\subsection{CES without energy (CES)}

@@

\subsection{CES with $Q$ (CES$q$)}

@@

\subsection{CES with $X$ (CES$x$)}

@@

\subsection{CES with $U$ (CES$u$)}

### 2013-04-14 EconGrowthSeminar.Rnw
Added the option to include CES models in partialResidualPlots.

\begin{frame}{Beneficial to Include Energy?}  %----------------------------
<<US_Partial_Residuals_Plot, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight>>=
partialResidualPlot(countryAbbrev="US", energyType="Q", textScaling=textScaleForGraphs)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{US}{0.055}
\end{frame}

\begin{frame}{Beneficial to Include Energy?}  %----------------------------
<<UK_Partial_Residuals_Plot, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight>>=
partialResidualPlot(countryAbbrev="UK", energyType="Q", textScaling=textScaleForGraphs)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{UK}{0.055}
\end{frame}

>>>>

\begin{frame}{Beneficial to Include Energy?}  %----------------------------
<<US_Partial_Residuals_Plot, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight, eval=TRUE>>=
partialResidualPlot(modelType="CD", countryAbbrev="US", energyType="Q", textScaling=textScaleForGraphs)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{US}{0.055}
\end{frame}

\begin{frame}{Beneficial to Include Energy?}  %----------------------------
<<UK_Partial_Residuals_Plot, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight>>=
partialResidualPlot(modelType="CD", countryAbbrev="UK", energyType="Q", textScaling=textScaleForGraphs)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{UK}{0.055}
\end{frame}
